### 2008-10-08 trunk/Papers/rmrs-08/conclusion.tex

In this paper, we have motivated and defined \rmrs---a semantic
framework that has been used to represent, compare, and combine
semantic information computed from deep and shallow parsers.  \rmrs\
is designed to be maximally flexible on the type of semantic
information that can be left underspecified, so that
the semantic output of a shallow parser needn't over-determine or
under-determine the semantics that can be extracted from
the shallow syntactic analysis.  This lays the formal foundations for
a formalism that is emerging as a standard in
robust semantic processing.

We then defined the concept of \emph{solved forms} of an \rmrs, in
analogy to dominance graphs.  Solved forms abstract away from
irrelevant differences between models, and have been a crucial concept
in the development of efficient solvers for dominance graphs.  In the
future, we hope to draw upon that body of literature to develop
efficient algorithms for processing \rmrs s, including solvers which
will compute fully specific semantic representations from an \rmrs,
efficient inference algorithms, and algorithms for entailment
checking.  We have argued that entailment is a central notion in
comparing the outputs of different {\sc nlp} tools, and provided a syntactic
characterisation for \rmrs\ entailment, which we hope will allow us to
develop efficient entailment algorithms in the future.

>>>>

In this paper, we have motivated and defined \rmrs---a semantic
framework that has been used to represent, compare, and combine
semantic information computed from deep and shallow parsers.  \rmrs\
is designed to be maximally flexible on the type of semantic
information that can be left underspecified, so that the semantic
output of a shallow parser needn't over-determine or under-determine
the semantics that can be extracted from the shallow syntactic
analysis.  The key contribution of this paper was to lay the formal
foundations for a formalism that is emerging as a standard in robust
semantic processing.

Although we have not directly provided new tools for modelling or
processing language, we believe that a cleanly defined model theory
for \rmrs\ is a crucial prerequisite for the future development of
such tools.  We hope that future research will build upon this paper
to develop efficient algorithms and implementations for solving \rmrs
s, performing inferences that enrich \rmrs s from shallow analyses
with deeper information, and checking consistency of \rmrs s that were
obtained from different parsers.

### 2008-10-08 trunk/Papers/rmrs-08/entailment.tex

\section{RMRS Comparison as Entailment}
\label{sec:entailment}

The definitions from Section~\ref{sec:rmrs} serve an important theoretical
purpose in that they provide a formal underpinning \rmrs's existing use
for partial semantic representations in implemented
systems.  Next to the peace of
mind that comes with the use of a formally understood formalism, a
clean definition can also provide a foundation for the development of
efficient algorithms; this has been shown in the past in the dominance
graph literature, most recently \newcite{KolRegTha08}.  We hope that
this paper can likewise provide a starting point for the development
of efficient solvers and inference mechanisms for \rmrs.

On the other hand, the above definition of \rmrs\ provides new
concepts which can help us phrase questions of practical grammar
engineering in well-defined formal terms.  For instance, one crucial
issue in developing a hybrid system that combines the outputs of deep
and shallow processors is to determine whether the \rmrs s produced by
the two systems are compatible.  Because we assume that the deep
grammar is more informative than that of the shallow processor, this
amounts to checking whether every piece of information in the shallow
representation is consistent with what the deep grammar said.  In
general, it isn't even obvious what ``consistency'' is supposed to
mean exactly, given that variables might be renamed, $\ARG$ roles
assigned in different ways, and structural relations underspecified.
But given the formal definition of \rmrs, we can now propose that an
\rmrs\ $\varphi'$ that was computed by a shallow processor is
consistent with an \rmrs\ $\varphi$ that was computed by a deep
grammar if $\varphi$ entails $\varphi'$ (i.e., $\varphi \models
\varphi'$).  If we could establish 
efficiently that $\varphi'$ is entailed by $\varphi$, this would tell
us that all models that satisfy $\varphi$ are also models of
$\varphi'$, i.e.\ the deep representation is formally more specific.

We cannot provide an efficient algorithm for testing entailment in
this paper.  However, we propose the following novel syntactic
characterisation of entailment.

\begin{definition}
  We call an \rmrs\ $\varphi'$ an \emph{extension} of an \rmrs\
  $\varphi$ iff \todo{fill this in -- don't have the def with me right
    now}.
\end{definition}

\begin{thm}\label{thm:big-one}
  Let $\varphi, \varphi'$ be two {\sc rmrs}s.  Then $\varphi \models
  \varphi'$ iff for every solved form $S$ of $\varphi$, there is a
  solved form $S'$ of $\varphi'$ such that $S$ is an extension of
  $S'$. 
\end{thm}
\begin{proof}[Proof (sketch)]
``$\Leftarrow$'' follows from
Props.~\ref{prop:solved-forms-are-satisfiable} and
\ref{prop:models-satisfy-solved-forms}.

``$\Rightarrow$'': The basic idea is to construct a solved form for
$\varphi'$ by choosing a solved form for $\varphi$ and removing all
atoms using variables that don't occur in $\varphi'$.  The hard part
is the proof that the result is a solved form of $\varphi'$; this step
involves proving that if $\varphi \models \varphi'$, then all {\sc ep}s in
$\varphi'$ also occur in $\varphi$.  
\end{proof}

This characterisation does not by itself provide an efficient test for
entailment because the number of solved forms of $\varphi$ and
$\varphi'$ may be exponential, and thus it is infeasible to compute
them all and then test for extensions.  Nevertheless, we hope that the
theorem provides a first step towards an improved understanding of
entailment that will lead to such an efficient algorithm in the
future.

>>>>

\section{Future work}
\label{sec:entailment}

The definitions from Section~\ref{sec:rmrs} serve an important
theoretical purpose in that they provide a formal underpinning \rmrs's
existing use for partial semantic representations in implemented
systems.  Next to the peace of mind that comes with the use of a
well-understood formalism, we hope that the work reported here will
serve as a starting point for future work on efficient solvers and
inference mechanisms.

One direction which we believe may develop from this paper is the
development of efficient solvers for \rmrs.  As a first step, it would
be interesting to define a practically useful fragment of \rmrs\ with
polynomial-time satisfiability.  Our definition is sufficiently close
to that of dominance constraints that we expect that it should be
feasible to carry over the definition of \emph{normal dominance
  constraints} \cite{Althaus_etal:JoA} to \rmrs; neither the lexical
ambiguity of the node labels nor the separate specification of
predicates and arguments should make satisfiability harder.

Furthermore, the above definition of \rmrs\ provides new concepts
which can help us phrase questions of practical grammar engineering in
well-defined formal terms.  For instance, one crucial issue in
developing a hybrid system that combines or compares the outputs of
deep and shallow processors is to determine whether the \rmrs s
produced by the two systems are compatible.  In the new formal terms
we can use now, we can characterise compatibility of a more detailed
\rmrs\ $\varphi$ (perhaps from a deep grammar) and a less detailed
\rmrs\ $\varphi'$ simply as \emph{entailment} $\varphi \models
\varphi'$.  If entailment holds, this tells us that all claims that
$\varphi'$ makes about the semantic content of a sentence is
consistent with the claims that $\varphi$ makes.

At this point, we cannot provide an efficient algorithm for testing
entailment of \rmrs.  However, we propose the following novel
syntactic characterisation which may again serve as a starting point
for research along those lines:

\todo{define ``extension''!!}

\begin{prop}\label{thm:big-one}
  Let $\varphi, \varphi'$ be two {\sc rmrs}s.  Then $\varphi \models
  \varphi'$ iff for every solved form $S$ of $\varphi$, there is a
  solved form $S'$ of $\varphi'$ such that $S$ is an extension of
  $S'$. 
\end{prop}

### 2008-10-08 trunk/Papers/rmrs-08/introduction.tex

Representing semantics as a logical form that supports automated
inference and model construction is vital for deeper language
engineering tasks, such as dialogue systems.  Logical forms
can be obtained from hand-crafted deep grammars
\cite{butt:etal:1999,copestake:flickinger:2000}, but this
tends to lack robustness: they fail to cover all
words and linguistic constructions and by design do not handle
ill-formed phrases. 
There has thus been a trend recently towards systems for
robust wide-coverage semantic construction (e.g., 
\cite{bos:etal:2004,zettlemoyer:collins:2007}).     
% COMMENT: this issue about optional arguments is really too cryptic.
% Basically, there is no way for assessing by *any* learned language
% processor whether a single word form that appears in
% a number of different syntactic valencies is an example of lexical
% sense ambiguity, or they are all examples of the same sense being
% used in syntactically different ways (i.e., we have optional
% arguments).  Only hand-crafted grammars can tell you this.  So, for
% example, Bos et al (2004) would get bet(x,y) for "I bet 5 pounds",
% and bet(x,y,z) for "I bet you 5 pounds" and bet(x,p) for "I bet it's
% going to rain": I.e., the vocab of the semantic formalism is
% constructed automatically from NL word forms, and the predicate
% symbols in that automatically constructed vocab vary in (a) valency,
% and (b) the semantic dependencies represented by a given argument
% position (e.g., the second argument of "bet" in "I bet 5 pounds" is
% 5 pounds, and for "I bet it's raining" it's "it's raining".
% 
But there are certain semantic phenomena that these robust
approaches don't capture reliably, including quantifier scope,
optional arguments, and long-distance dependencies (for instance,
\newcite{clark:etal:2004} report that the parser used by
\newcite{bos:etal:2004} yields 63\% accuracy on object
extraction; e.g., {\em the man that I met\dots}).  Forcing a robust
parser to make a decision about these phenomena can therefore be
error-prone.  Depending on the application, it may be preferable to
give the parser the option to leave a semantic decision open when it's not
sufficiently informed---i.e., to compute a partial semantic
representation and to complete it later, using information
extraneous to the parser.

In this paper, we focus on an approach to representing partial
semantic information that supports this strategy: Robust Minimal
Recursion Semantics (\rmrs, \cite{copestake:2007a,copestake:2007b}).
\rmrs\ is 
designed to
support underspecification of scope and predicate-argument structure.
It is an emerging standard for representing partial
semantic information, and has been applied in several implemented
systems.  For instance, \newcite{copestake:2003} and
\newcite{frank:2004} use it to specify semantic components to shallow
parsers ranging in depth from {\sc pos} taggers to chunk parsers and
intermediate parsers such as {\sc rasp} \cite{briscoe:etal:2006}.
{\sc mrs} analyses \cite{copestake:etal:2005}
derived from deep grammars, such as the English
Resource Grammar ({\sc erg}, \cite{copestake:flickinger:2000}) are
special cases of {\sc rmrs}.

However, from a formal perspective, \rmrs\ is somewhat ad-hoc in that
it has no clearly defined model theory; i.e., what exactly any given
\rmrs\ representation actually \emph{means} is not defined.  The key
contribution we make is to cast {\sc rmrs}, for the first time, as a
logic with a well-defined model theory.  This has a number of
advantages.  First, it allows us to formally relate \rmrs\ to other
frameworks such as \mrs\ and predicate logic in order to establish the
compatibility of \rmrs\ with the approaches to robust semantic
analysis mentioned above.  
Second, it provides us with a notion of logical
entailment, which we can use to verify the compatibility of \rmrs\ 
representations computed from different parsers, such as a deep and a
shallow parser.  Finally, it is only through rigid formalisation that
we can give users of \rmrs\ access to inference systems and to
algorithms for efficiently computing the possible fully specific
semantic representations described by the \rmrs\ output of a shallow
parser.

Section~\ref{sec:motivation} illustrates via example the differences
and challenges with deep and shallow semantic construction.
Section~\ref{sec:rmrs} defines the syntax and model-theory of \rmrs.
In Section~\ref{sec:entailment} we demonstrate the practical
usefulness of our formalisation by modelling compatibility of partial
semantic representations as \rmrs\ entailment and by providing a
syntactic criterion for characterising entailment.
Section~\ref{sec:conclusion} concludes and points to future work.

>>>>

Representing semantics as a logical form that supports automated
inference and model construction is vital for deeper language
engineering tasks, such as dialogue systems.  Logical forms
can be obtained from hand-crafted deep grammars
\cite{butt:etal:1999,copestake:flickinger:2000}, but this
tends to lack robustness: they fail to cover all
words and linguistic constructions and by design do not handle
ill-formed phrases. 
There has thus been a trend recently towards systems for
robust wide-coverage semantic construction (e.g., 
\cite{bos:etal:2004,zettlemoyer:collins:2007}).     
% COMMENT: this issue about optional arguments is really too cryptic.
% Basically, there is no way for assessing by *any* learned language
% processor whether a single word form that appears in
% a number of different syntactic valencies is an example of lexical
% sense ambiguity, or they are all examples of the same sense being
% used in syntactically different ways (i.e., we have optional
% arguments).  Only hand-crafted grammars can tell you this.  So, for
% example, Bos et al (2004) would get bet(x,y) for "I bet 5 pounds",
% and bet(x,y,z) for "I bet you 5 pounds" and bet(x,p) for "I bet it's
% going to rain": I.e., the vocab of the semantic formalism is
% constructed automatically from NL word forms, and the predicate
% symbols in that automatically constructed vocab vary in (a) valency,
% and (b) the semantic dependencies represented by a given argument
% position (e.g., the second argument of "bet" in "I bet 5 pounds" is
% 5 pounds, and for "I bet it's raining" it's "it's raining".
%

But there are certain semantic phenomena that these robust
approaches don't capture reliably, including quantifier scope,
optional arguments, and long-distance dependencies (for instance,
\newcite{clark:etal:2004} report that the parser used by
\newcite{bos:etal:2004} yields 63\% accuracy on object
extraction; e.g., {\em the man that I met\dots}).  Forcing a robust
parser to make a decision about these phenomena can therefore be
error-prone.  Depending on the application, it may be preferable to
give the parser the option to leave a semantic decision open when it's not
sufficiently informed---i.e., to compute a partial semantic
representation and to complete it later, using information
extraneous to the parser.

In this paper, we focus on an approach to representing partial
semantic information that supports this strategy: Robust Minimal
Recursion Semantics (\rmrs, \cite{copestake:2007a,copestake:2007b}).
\rmrs\ is designed to support underspecification of lexical ambiguity,
scope, and predicate-argument structure.  It is an emerging standard
for representing partial semantic information, and has been applied in
several implemented systems.  For instance, \newcite{copestake:2003}
and \newcite{frank:2004} use it to specify semantic components to
shallow parsers ranging in depth from {\sc pos} taggers to chunk
parsers and intermediate parsers such as {\sc rasp}
\cite{briscoe:etal:2006}.  {\sc mrs} analyses
\cite{copestake:etal:2005} derived from deep grammars, such as the
English Resource Grammar ({\sc erg}, \cite{copestake:flickinger:2000})
are special cases of {\sc rmrs}.

The key contribution we make is to cast {\sc rmrs}, for the first
time, as a logic with a well-defined model theory.  Previously, no
such model theory existed, and so \rmrs\ had to be used in a somewhat
ad-hoc manner which left it open what exactly any given \rmrs\
representation actually \emph{means}.  This has hindered quite
practical progress in terms of understanding the relationship of
\rmrs\ to other frameworks such as \mrs\ and predicate logic and in
terms of the development of efficient algorithms.  As one application
of the formal concepts that come with our definitions, we propose a
novel way of characterising consistency of \rmrs\ analyses across
different parsers as entailment.

Section~\ref{sec:motivation} introduces \rmrs\ informally and
illustrates why it is necessary and useful for representing semantic
information across deep and shallow language processors.
Section~\ref{sec:rmrs} defines the syntax and model-theory of \rmrs.
We finish in Section~\ref{sec:entailment} by pointing out some avenues
for future research, and conclude in Section~\ref{sec:conclusion}.
